Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-02-05 paper/paper.tex
even more updates to methods and pill section of paper Tested everything on node32.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}\fixme{whats the 'sim'
    here?}
\end{table}

@@

At $t=0$ in Figs.~\ref{image-p}, the cytoplasmic MinE
is concentrated in the lower portion of the cell, and is diffusing
upward.  The MinD protein is initially concentrated on the membrane in
the upper portion of the cell.  At the boundary between the
cytoplasmic MinE and the membrane-bound MinD, there is a ring of the
membrane-bound MinE:MinD complex, which is moving towards the upper
end of the cell, removing the MinD from the membrane as it goes.  The
MinD is removed from the membrane in its MinD:ADP form, which begins
to accumulate at the upper end of the cell, peaking at around
10 seconds.

At 15 seconds, the membrane-bound MinD:ATP has been essentially
removed from the top end of the cell, and MinD:ATP has begun to bind
to the membrane in the the lower half of the cell.  At this stage,
there is a high concentration of cytoplasmic MinE at the top of the
cell, and by 16 seconds we begin to see the formation of a MinE ring
just below the center of the cell.  At 18 seconds, the cell has
reached its original state, with MinD:ATP bound to the membrane on the
lower third of the cell, a high concentration of cytoplasmic MinE in
the upper half of the cell, and a MinE ring pushing on the
membrane-bound MinD:ATP.

The reactions shown in Figure~\ref{fig:reactions} are cyclical so we
can begin our observations at any stage in the cycle (which is
presented for the natural, pill shapes cell in Fig.~\ref{image-p}
below).  We consider first MinD:ATP floating freely in the cytoplasm.
The protein will not react with any other floating protein, and it
will not spontenously shift back into the MinD:ADP stage of the cycle.
It will diffuse until it attaches to a cell membrane.  This will
happen most rapidly in cytoplasmic regions that are close to membrane
regions of high curvature, where diffusion in a random direction is
more likely to lead to an encounter with a wall.  This effect is made
more pronounced by the MinD:ATP tendency to attract and cluster on the
membrane, as reflected in Equation~\ref{eq:d-on-wall}, which shows
that when adjacent to the membrane, the rate of binding to the
membrane is porportional to the amount of MinD already bound to the
wall.  Initially, then, there is a strong tendency of MinD to cluster
on the polar regions of the cell, where the curvature is highest.

We next see the formation of MinE `rings' on the membranes of the
polar caps.  As the freely diffusing MinE enters the cap portion it
begins to encounter concentrated membrane bound MinD.  The MinE binds
to this MinD and remains for a time on the walls in compound form.
When released, if it diffuses inward, there is less cross sectional
wall for it to bump into and also less concentrated MinD:ATP to react
with, whereas if it diffuses outward the opposite is the case.  This
process results in the observed qualitative 'encroaching MinE ring'
behavoir that releases MinD:ADP back into the cytoplasm.  This
MinD:ADP will diffuse for a time before spontanosly undergowing
nucleotide exchange and reaching the MinD:ATP state that can bind to
the membrane again.  There is a strong chance that it has by this time
diffused out of the polar region and is on its way to the opposite
polar region of the cell.  If it hasn't then it will reattach to the
MinD:ATP on the cell wall deeper into the endcap and will eventually
be converged upon by the MinE ring, only to be released again in
MinD:ADP form.

Overall this process results in extremely robust oscillatory
behavoir of polar selection and oscillation in a wide variety of cell
shapes.

%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-300-50-0-0-1500}
%%   \caption{Total protein fluctuation in the right, middle, and left
%%     parts of a natural pill shaped cell.  The cell is the same as is
%%     shown in Fig.~\ref{image-p}.  The time axis is also identical to
%%     that shown in Fig.~\ref{image-p}, and shows one period of the
%%     oscillation after the system has settled down.  The vertical axis
%%     shows stacked the total number of proteins that are of four
%%     different compound stages and in the right section of the cell
%%     (bottom four colors), in the middle of the cell (middle four
%%     colors) and in the left section of the cell (top four colors). The
%%     different forms of protein changes while the total number of MinD
%%     protein in the cell remains constant, shown by the fact that the
%%     very top line is horizontal.}
%%   \label{box-p}
%% \end{figure}

>>>>

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}\fixme{the 'sim' here
    means that it's waiting for a simulation.  Not sure where it went
    but I'm rerunning it.}
\end{table}

@@

At $t=0$ in Figs.~\ref{image-p}, the cytoplasmic MinE is concentrated
in the lower portion of the cell and is diffusing upward, since it
will react with and stick to wall-bound MinD that is concentrated on
the membrane in the upper portion of the cell.  At the boundary
between the cytoplasmic MinE and the membrane-bound MinD, there is a
ring of the membrane-bound MinE:MinD complex, which is moving towards
the upper end of the cell, removing the MinD from the membrane as it
goes.  The MinD is removed from the membrane in its MinD:ADP form and
diffuses freely for a time before spontenously undergowing nucleotide
exchange and shifting back into the MinD:ATP that is able to attach to
the membrane.  Diffusion away from the end of the cell will bring the
protein into the MinE ring, where reactions will prevent it from
collecting for any significant time on the membrane.  Diffusion
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, peaking at around 10 seconds in Figure~\ref{image-p}.  As
the MinE ring approaches the end of the cell, seen in the seconds
following 10, the only freely diffusing MinD:ADP that 'escapes' the
MinE will be that which has diffused away from the the end of the cell
and past the MinE ring towards the center.

At 15 seconds, the membrane-bound MinD:ATP has been essentially
removed from the top end of the cell, and MinD:ATP has begun to bind
to the membrane in the the lower half of the cell.  At this stage,
there is a high concentration of cytoplasmic MinE at the top of the
cell, and by 16 seconds we begin to see the formation of a MinE ring
just below the center of the cell.  At 18 seconds, the cell has
reached its original state (reversed directionally), with MinD:ATP
bound to the membrane on the lower third of the cell, a high
concentration of cytoplasmic MinE in the upper half of the cell, and a
MinE ring pushing downward on the membrane-bound MinD:ATP.

In general process results in extremely robust oscillatory behavoir of
polar selection and oscillation in a wide variety of cell shapes.
